
```{r }
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
```{r }

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results = TRUE)
library(dplyr)
library(lubridate)
library(chron)
library(ggplot2)
```

## Loading and preprocessing the data

```{r 1. Code for reading in the dataset and/or processing the data}
dataFileName<-"activity.csv"
zipFileName<-"Activity monitoring data.zip"
fileUrl<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if (!(file.exists(dataFileName))){
        download.file(fileUrl,zipFileName)
        unzip(zipFileName,exdir=getwd())
    }
if (file.exists(zipFileName)) file.remove(zipFileName)
activity<-read.csv(dataFileName)
```

## What is mean total number of steps taken per day?

### Daily Step Count Histogram

```{r 2. Histogram of the total number of steps taken each day}
knitr::fig_path('figures/')
dailySteps<-aggregate(activity[,1],list(activity$date),sum)
hist(dailySteps$x,xlab="Daily Step Cout",main="")
```

### Daily Steps Mean and Median

```{r 3. Mean and median number of steps taken each day}
c(mean=mean(dailySteps$x,na.rm=TRUE), median=median(dailySteps$x,na.rm=TRUE))
```

## What is the average daily activity pattern?

### Average Steps Taken at Each Time Interval
```{r 4. Time series plot of the average number of steps taken}
IntervalAverageSteps<-aggregate(activity[,1],list(activity$interval),FUN=mean,
                               na.rm=TRUE, na.action=NULL)
names(IntervalAverageSteps)<-c("Interval","Average Steps")
plot(IntervalAverageSteps$Interval, IntervalAverageSteps$`Average Steps`,type="l")
```

### Most Active Iterval

```{r 5. The 5-minute interval that, on average, contains the maximum number of steps}
IntervalAverageSteps[IntervalAverageSteps$`Average Steps`==max(IntervalAverageSteps$`Average Steps`),1]
```

## Inputing missing values

```{r 6. Code to describe and show a strategy for inputing missing data}
#replace NA
activity<-split(activity,is.na(activity$steps))
matchIntervals<-match(activity$`TRUE`$interval,IntervalAverageSteps$Interval)
activity$`TRUE`$steps<-IntervalAverageSteps$`Average Steps`[matchIntervals]
activity<-rbind(activity$`FALSE`,activity$`TRUE`)
```

### Daily Step Count Histogram After Replacing Missing Values for averages

```{r 7. Histogram of the total number of steps taken each day after missing values are imputed}
dailySteps<-aggregate(activity[,1],list(activity$date),sum)
hist(dailySteps$x)
```

### Daily Steps Mean and Median After Replacing Missing Values for averages

```{r 7.1 Mean and median number of steps taken each day after missing values are imputed}
c(mean=mean(dailySteps$x,na.rm=TRUE), median=median(dailySteps$x,na.rm=TRUE))
```

After repacing the missing values for the mean the mean value does not change, as expected. The median however ir closer to the mean now.

## Are there differences in activity patterns between weekdays and weekends?

```{r 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends}

# Add column to determine weekday

activity<-cbind(activity,weekend=is.weekend(activity$date))


#Calculate averages for each interval and each weekday type

IntervalWeekdayMeanSteps<-aggregate(activity[,1],list(activity$interval,activity$weekend),FUN=mean,
                               na.rm=TRUE, na.action=NULL)
names(IntervalWeekdayMeanSteps)<-c("Interval","Weekday","Steps")
IntervalWeekdayMeanSteps$Weekday<-as.factor(IntervalWeekdayMeanSteps$Weekday)
levels(IntervalWeekdayMeanSteps$Weekday)<-c("Weekday","Weekend")
names(IntervalWeekdayMeanSteps)<-c("Interval","Weekday","Steps")

# Plot the result

qplot(Interval,Steps, data=IntervalWeekdayMeanSteps,color=Weekday)+stat_smooth(aes(x = Interval, y = Steps), method = "lm",
              formula = y ~ poly(x, 20), se = FALSE)

```
```
```

